Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-07-27 manuscript/usses_water_ms.tex
code and paper updates

% 
\begin{abstract}
Precipitation is predicted to become more variable in the western United
States, meaning years of above and below average precipitation will
become more common. Periods of drought and deluge are major drivers of
ecosystem functioning in water limited grasslands, but how ecosystems
respond to precipitation may change as the duration of above and below
average periods increases. Changes in ecosystem functional response
could reflect compensatory changes in species composition or species
reaching physiological thresholds at extreme precipitation levels. Here
we report the results of a five-year experiment where we used drought
and irrigation treatments (approximately 50\% decrease/increase) to
investigate whether ecosystem functional response remains consistent at
extreme precipitation levels. We also use species-level adundance data
to link ecosystem functional response to community dynamics. Ecosystem
functional response, defined as the relationship between soil moisture
and aboveground net primary productivity, was suprisingly consistent
across drought, control, and irrigation treatments. Aboveground net
primary productivity (ANPP) responded positively to availabile soil
moisture, but the response was weak and did not vary across treatments.
There was also no evidence that treatment effects grew over time. The
similarity of ecosystem functional response across treatments was not
due to compensatory shifts at the plant community level, where species
composition among treatments was similar and remarkably stable over the
five years. At least in the short-term, ecosystem functional response
and community composition in this sagebrush steppe system is consistent,
even at extreme precipitation levels. Such consistency may be due to
bet-hedging strategies of constituent species.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

@@

Ecosystems worldwide are projected to experience either wetter or drier
conditions in the future \citep{Zhang2007, Greve2014}. This has
important implications for ecosystem functioning because mean annual net
primary productivity (ANPP) positively covaries with mean annual
precipitation across ecosystems \citep{Sala1988, Huxman2004} and
interannual responses of ANPP are linked to interannual variability in
precipitation within ecosystems
\citep{Lauenroth1992, Hsu2012, Gherardi2015a}. To predict how ecosystems
might respond to extreme precipitation levels, we need to understand
whether ecosystems tend to follow temporal trends, which assume
relatively constant plant community composition, or spatial trends,
which allow community composition to shift \citep{Wilcox2016}.

At a given site, the functional response of ANPP to water availability
(e.g., soil moisture) can be characterized by fitting a model to
historical observations of ANPP and soil moisture. However, the fitted
functional response may provide an incomplete picture because future
conditions are likely to be outside the historical range of variability
\citep{Smith2011}. For example, historical trends may underestimate the
potential for the soil moisture-ANPP relationship to saturate if soil
moisture is pushed far beyond typical levels. Saturating relationships
are actually common \citep{Hsu2012, Gherardi2015a}, perhaps because
other resources, like nitrogen, become more limiting in wet years than
dry years. Knowing the curvature of the soil moisture-ANPP relationship
at extreme precipitation levels is critical for understanding how
ecosystems will respond to chronic alterations in water availability.

Another problem with relying on historical ecoystem functional responses
is that they are not static. Changes in species identities and
abundances can alter an ecosystem's functional response to water
availability because different species have different physiological
thresholds for producting biomass. \citet{Smith2009} introduced the
`Hierarchical Response Framework' (HRF) for understanding the interplay
of community composition and ecosystem functioning in response to
resource manipulations over time. In the near term, ecosystem
functioning such as annual net primary productivity (ANPP) will reflect
the physiological responses of individual species to the manipulated
resource level. For example, ANPP may decline under simulated drought
because the initial community consisted of drought-intolerant species
\citep{Hoover2014}. Over longer time spans, ecosystem functioning may
recover as new species colonize or initial species reorder in relative
abundance. For example, ANPP may initially decline, but eventually rise
back to pre-treatment levels once drought-tolerant species become more
abundant and compensate for drought-intolerant species
\citep{Hoover2014}. It is also possible that ecosystem functioning
shifts to a new mean state, reflecting the suite of species in the new
community \citep{Knapp2012}.

@@

To that end, here we report the results of a five-year precipitation
manipulation experiment in a sagebrush steppe grassland. We imposed
drought and irrigation treatments (approximately \(\pm50\%\)) and
measured ecosystem (ANPP) and community (species composition) responses.
We focus on how the drought and irrigation treatments affect the
relationship between available soil moisture and ANPP, and if community
dynamics underly the ecosystem responses. In particular, we are
interested in the consistency of the soil moisture-ANPP relationship
among treatments. Is the relationship steeper under the drought
treatment, at low soil moisture? Does the relationship saturate under
the irrigation treatment, at high soil moisture? To answer these
questions we fit a random intercepts, random slopes model to test
whether the regressions differed among treatments. We also analyzed
community composition over time, allowing us to place our experimental
results within the framework of the HRF and our competing predictions
(Fig. 1).

@@

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the United States Sheep Experimental Station (USSES)
near Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m
above sea level. The plant community is dominated by the shrub
\emph{Artemesia tripartita} and three perennial bunchgrasses,
\emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and
\emph{Hesperostipa comata}. During the period of our experiment (2011 --
2015), average mean annual precipitation was 265 mm
year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July.

@@

Between 1926 and 1932, range scientists at the USSES established 26
permanent 1 m\(^2\) quadrats to track vegetation change over time. In
2007, we (well, one of us {[}P. Adler{]}) relocated 14 of the original
quadrats, six of which were inside a large, permanent livestock
exclosure. We use these six plots as control plots that have recieved no
treatment, just ambient precipitation. In spring 2011, we (well, two of
us {[}A. Kleinhesselink and P. Adler{]}) established 16 new 1 m\(^2\)
plots located in the same exclosure as the six control plots. We avoided
areas on steep hill slopes, areas with greater than 20\% cover of bare
rock, and areas with greater than 10\% cover of the shrubs \emph{Purshia
tridentata} and/or \emph{Amelanchier utahensis}. We established the new
plots in pairs and randomly assigned each plot in a pair to receive a
``drought'' or ``irrigation'' treatment.

@@

We then modeled the daily deviation from ambient conditions of the
drought and irrigation treatments using a linear mixed effects model
with independent variables for treatment, season (winter, spring,
summer, fall), rainfall, and all two-way interactions. Rainy days were
defined as any day in which precipitation was recorded and average
temperature was above 3\(^{\circ}\)C. The day immediately following
rainfall was also classified as rainy. We fit the model using the
`lme4::lmer()' function \citep{Bates2015} in R \citep{R2016}, with
random effects for plot group and date. We weighted observations by the
number of unique sensors or spot measurements that were taken in each
plot on that day. We then used the model to predict the average daily
soil moisture in the treated plots based on the average daily ambient
soil moisture. We could only predict soil moisture in the treated plots
on days for which we took at least one ambient soil moisture measurement
(Fig. 2B).

@@

Species composition data came from two sources: yearly census maps for
each plot made using a pantograph \citep{Hill1920} and yearly counts of
annual species in each plot. From those maps, we have data on the
density of all annuals and perennials forbs, basal cover of perennial
grasses, and canopy cover of shrubs. We made a large plot-treatment-year
by species matrix, where columns were filled with either basal cover or
density, depending on the measurement made for the particular species.
So we could analyze the different types of data together, we
standardized the values in each column. This puts all abundance values
on the same scale, meaning that common and rare species are weighted
equally. Nonetheless, if we assume that rare species will respond more
than common ones, then this approach is anti-conservative for detecting
community change. This means that our approach is biased toward
detecting compositional changes.\todo{Did I say this right?}

@@

Our main goal was to test whether the relationship between ANPP and soil
moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site
\citep{Blaisdell1958, Dalgleish2011, Adler2012}, we chose to use
cumulative volumetric water content from March through June as our
metric of soil moisture (hereafter referred to simply as `soil
moisture'). To achieve this goal, we fit a multi-level random intercept
and random slope regression with log(ANPP) as the response variabile and
soil moisture as the sole predictor. We log-transformed ANPP to account
for heteroscedasticity. We fit the model under a Bayesian framework,
allowing us to test for treatment differences by comparing the posterior
distributions of the treatment-level coefficients
\citep[e.g.,][]{Tredennick2013}. Both log(ANPP) and soil moisture were
standardized to have mean 0 and unit variance before fitting the model
{[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.

@@

The intercept and slope terms are modeled hierarchichally to account for
the non-independence of observations across years within plots and to
allow us to test the hypothesis that our treatments alter the ANPP-soil
moisture relationship. Thus, our multi-level model has three grouping
levels for coefficients, representing the nested structure of the data:
(i) overall coefficients, (ii) treatment coefficients, and (iii) plot
coefficients. Each subsequent level is drawn from the distribution of
coefficients at the previous level, meaning plot-level coefficients are
drawn from treatment-level coefficients, which are drawn from overall
coefficients. We also include a covariance structure among the intercept
and slope at each level. Formally, our hierarhical structure is as
follows, where we drop the intercept (0) and slope (1) subscripts and
instead refer to a vector of coefficients, \(\boldsymbol{\beta}\):

@@

We fit the model using a Bayesian approach, obtaining posterior
estimates of all unkowns via the No-U-Turn Hamiltonian Monte Carlo
sampler in Stan \citep{stan2016}. We used the R package `rstan'
\citep{rstan2016} to link R \citep{R2016} to Stan. We obtained samples
from the posterior distribution for all model parameters from four
parallel MCMC chains run for 10,000 iterations, saving every
\(10^{\text{th}}\) sample. Traceplots of all parameters were visually
inspected to ensure well-mixed chains and convergence. We also made sure
all scale reduction factors (\(\hat{R}\)) were less than 1.1.

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP
(mean of \(\beta_{1}\) = 0.41; Fig. 3), but the effect is associated
with high uncertainty, with credible intervals that overlap zerp (80\%
BCI = -0.08, 0.86; 95\% BCI = -0.45, 1.13). Ecosystem functional
response was similar among treatments, with treatment level regressions
between soil moisture and log(ANPP) having similar intercepts and slopes
(Fig. 3A). There was also no evidence that the treatment effects became
more important over time because there was no directional trend in the
random year effects (Fig. 4).

@@

The similarity of ecosystem functional response (Fig. 3A) and community
composition (Fig. 5) across treatments is surprising because grasslands
generally, and sagebrush steppe specifically, are considered
water-limited systems. Indeed, we expected ecosystem functional
response, community composition, or both to change with precipitation
treatment, landing us in any box of Fig. 1 \emph{except} the top left.
So, why did our treatments fail to induce ecosystem or community
responses? We can think of three reasons; two are limitations of our
study, and one is the life history traits of the species in our focal
communities. We first discuss the potential limitations of our study,
and then discuss a biological explanation.

First, it could be that our manipulations were not large enough to
induce a response. That is, maybe a 50\% decrease/increase in any given
year is not abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our manipulations \emph{should} have been
large enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation neccesary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site. The 1\% quantile of
precipitation at our site is 110 mm, a 47\% reduction from the mean, and
the 99\% quantile is 414 mm, a 77\% increase from mean growing season
precipitation (Appendix 3). Thus, our drought treatment represented
extreme precipitation amounts, especially in years where ambient
precipitation was below average (Fig. 2A). The irrigation treatment may
have been too small, however.

Second, ANPP at our site may be influenced by factors beyond the window
of soil moisture we included in our statistical model. For example,
temperature can impact ANPP directly \citep{Epstein1997} and by
exacerbating the effects of soil moisture \citep{DeBoeck2011}.
Measurements of soil moisture likely contain a signal of temperature,
through its impact on evaporation and infiltration, but the measurements
will not capture the direct effect of temperature on metabolic and
physiological processes. Likewise, we did not redistribute snow across
our treatments in the winter, and snowmelt may spur early spring growth.
These statistical issues of missing potentially important covariates
could explain the weak and uncertain relationship we observed between
soil moisture and ANPP.

@@

NSF DBI-1400370 to Andrew T. Tredennick.\\
NSF Graduate Research Fellowship to Andrew R. Kleinhesselink.\\
NSF DEB-1353078 and DEB-1054040 to Peter B. Adler.

@@

\begin{itemize}
  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed
drafts of the paper.
  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
  \item Bret Taylor...
  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed
drafts of the paper.
\end{itemize}

@@

\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis` tests whether treatments form unique clusters in multidimensial space; `betadisper` tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  21 &   2 & 1.02 & 0.45 \\ 
  2011 & betadisper &  21 &   2 & 2.23 & 0.14 \\ 
  2012 & adonis &  22 &   2 & 1.10 & 0.29 \\ 
  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
  2013 & adonis &  22 &   2 & 1.23 & 0.14 \\ 
  2013 & betadisper &  22 &   2 & 0.28 & 0.76 \\ 
  2014 & adonis &  22 &   2 & 0.95 & 0.58 \\ 
  2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
  2015 & adonis &  21 &   2 & 1.05 & 0.37 \\ 
  2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
  2016 & adonis &  21 &   2 & 1.07 & 0.35 \\ 
  2016 & betadisper &  21 &   2 & 0.50 & 0.62 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}

>>>>

% 
\begin{abstract}
\textbf{Background.} Precipitation is predicted to become more variable
in the western United States, meaning years of above and below average
precipitation will become more common. Periods of extreme precipitation
are major drivers of ecosystem functioning in water limited grasslands,
but how ecosystems respond to precipitation may change as the duration
of above and below average periods increases. Changes in ecosystem
functional response could reflect compensatory changes in species
composition or species reaching physiological thresholds at extreme
precipitation levels.

\textbf{Methods.} We conducted a five year precipitation manipulation
experiment in a sagebrush steppe ecosystem in Idaho, US. We used drought
and irrigation treatments (approximately 50\% decrease/increase) to
investigate whether ecosystem functional response remains consistent at
extreme precipitation levels. We recorded data on aboveground net
primary productivity (ANPP), species abundance, and soil moisture. We
fit a random intercept, random slope model to determine if the
relationship between ANPP and soil moisture differed among treatments.
We used nondimensional multivariate scaling to quantify community
composition over the five years.

\textbf{Results.} Ecosystem functional response, defined as the
relationship between soil moisture and ANPP, did not vary across
drought, control, and irrigation treatments. ANPP responded positively
to availabile soil moisture, but the response was weak and uncertain.
There was also no evidence that treatment effects grew over time. Plant
community composition was remarkably stable over the course of the
experiment and did not differ among treatments.

\textbf{Discussion.} At least in the short-term, ecosystem functional
response and community composition in this sagebrush steppe system is
consistent, even at extreme precipitation levels. The similarity of
ecosystem functional response across treatments was not due to
compensatory shifts at the plant community level, but instead reflects
the consistent responses of individual species. Such consistency may be
due to bet-hedging strategies where species respond weakly to soil
moisture to ensure low variance of long-term success.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

@@

At a given site, the functional response of aboveground net primary
productivity (ANPP) to water availability (e.g., soil moisture) can be
characterized by fitting a model to historical observations of ANPP and
soil moisture. However, the fitted functional response may provide an
incomplete picture because future conditions are likely to be outside
the historical range of variability \citep{Smith2011}. For example,
historical trends may underestimate the potential for the soil
moisture-ANPP relationship to saturate if soil moisture is pushed far
beyond typical levels. Saturating relationships are actually common
\citep{Hsu2012, Gherardi2015a}, perhaps because other resources, like
nitrogen, become more limiting in wet years than dry years. Knowing the
curvature of the soil moisture-ANPP relationship at extreme
precipitation levels is critical for understanding how ecosystems will
respond to chronic alterations in water availability.

Another problem with relying on historical ecosystem functional
responses is that they are not static. Changes in species identities and
abundances can alter an ecosystem's functional response to water
availability because different species have different physiological
thresholds for producing biomass. \citet{Smith2009} introduced the
`Hierarchical Response Framework' (HRF) for understanding the interplay
of community composition and ecosystem functioning in response to
resource manipulations over time. In the near term, ecosystem
functioning such as ANPP will reflect the physiological responses of
individual species to the manipulated resource level. For example, ANPP
may decline under simulated drought because the initial community
consisted of drought-intolerant species \citep{Hoover2014}. Over longer
time spans, ecosystem functioning may recover as new species colonize or
initial species reorder in relative abundance. For example, ANPP may
initially decline, but eventually rise back to pre-treatment levels once
drought-tolerant species become more abundant and compensate for
drought-intolerant species \citep{Hoover2014}. It is also possible that
ecosystem functioning shifts to a new mean state, reflecting the suite
of species in the new community \citep{Knapp2012}.

@@

To that end, here we report the results of a five-year precipitation
manipulation experiment in a sagebrush steppe grassland. We imposed
drought and irrigation treatments (approximately \(\pm50\%\)) and
measured ecosystem (ANPP) and community (species composition) responses.
We focus on how the drought and irrigation treatments affect the
relationship between available soil moisture and ANPP, and if community
dynamics underlie the ecosystem responses. In particular, we are
interested in the consistency of the soil moisture-ANPP relationship
among treatments. Is the relationship steeper under the drought
treatment, at low soil moisture? Does the relationship saturate under
the irrigation treatment, at high soil moisture? To answer these
questions we fit a random intercept, random slope model to test whether
the regressions differed among treatments. We also analyzed community
composition over time, allowing us to place our experimental results
within the framework of the HRF and our competing predictions (Fig. 1).

@@

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the United States Sheep Experimental Station (USSES)
near Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m
above sea level. The plant community is dominated by the shrub
\emph{Artemesia tripartita} and three perennial bunchgrasses,
\emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and
\emph{Hesperostipa comata}. During the period of our experiment (2011 --
2015), average mean annual precipitation was 265 mm
year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
and 1932, range scientists at the USSES established 26 permanent 1
m\(^2\) quadrats to track vegetation change over time. In 2007, we
(well, one of us {[}P. Adler{]}) relocated 14 of the original quadrats,
six of which were inside a large, permanent livestock exclosure. We use
these six plots as control plots that have received no treatment, just
ambient precipitation, in the experiment described below.

@@

In spring 2011, we (well, two of us {[}A. Kleinhesselink and P.
Adler{]}) established 16 new 1 m\(^2\) plots located in the same
exclosure as the six control plots. We avoided areas on steep hill
slopes, areas with greater than 20\% cover of bare rock, and areas with
greater than 10\% cover of the shrubs \emph{Purshia tridentata} and/or
\emph{Amelanchier utahensis}. We established the new plots in pairs and
randomly assigned each plot in a pair to receive a ``drought'' or
``irrigation'' treatment.

@@

We then modeled the daily deviation from ambient conditions of the
drought and irrigation treatments using a linear mixed effects model
with independent variables for treatment, season (winter, spring,
summer, fall), rainfall, and all two-way interactions. Rainy days were
defined as any day in which precipitation was recorded and average
temperature was above 3\(^{\circ}\)C. The day immediately following
rainfall was also classified as rainy. We fit the model using the
`lme4::lmer()' function \citep{Bates2015} in R \citep{R2016}, with
random effects for plot group and date. We weighted observations by the
number of unique sensors or spot measurements that were taken in each
plot on that day. We then used the model to predict the average daily
soil moisture in the treated plots based on the average daily ambient
soil moisture. We could only predict soil moisture in the treated plots
on days for which we took at least one ambient soil moisture
measurement.

@@

Species composition data came from two sources: yearly census maps for
each plot made using a pantograph \citep{Hill1920} and yearly counts of
annual species in each plot. From those maps, we have data on the
density of all annuals and perennials forbs, basal cover of perennial
grasses, and canopy cover of shrubs. We made a large plot-treatment-year
by species matrix, where columns were filled with either basal cover or
density, depending on the measurement made for the particular species.
So we could analyze the different types of data together, we
standardized the values in each column. This puts all abundance values
on the same scale, meaning that common and rare species are weighted
equally. Nonetheless, if we assume that rare species will respond more
than common ones, then this approach is anti-conservative for detecting
community change. This means that our approach is biased toward
detecting compositional changes.

@@

Our main goal was to test whether the relationship between ANPP and soil
moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site
\citep{Blaisdell1958, Dalgleish2011, Adler2012}, we chose to use
cumulative volumetric water content from March through June as our
metric of soil moisture (hereafter referred to simply as `soil
moisture'). To achieve this goal, we fit a multi-level random intercept
and random slope regression with log(ANPP) as the response variable and
soil moisture as the sole predictor. We log-transformed ANPP to account
for heteroscedasticity. We fit the model under a Bayesian framework,
allowing us to test for treatment differences by comparing the posterior
distributions of the treatment-level coefficients
\citep[e.g.,][]{Tredennick2013}. Both log(ANPP) and soil moisture were
standardized to have mean 0 and unit variance before fitting the model
{[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.

@@

The intercept and slope terms are modeled hierarchically to account for
the non-independence of observations across years within plots and to
allow us to test the hypothesis that our treatments alter the ANPP-soil
moisture relationship. Thus, our multi-level model has three grouping
levels for coefficients, representing the nested structure of the data:
(i) overall coefficients, (ii) treatment coefficients, and (iii) plot
coefficients. Each subsequent level is drawn from the distribution of
coefficients at the previous level, meaning plot-level coefficients are
drawn from treatment-level coefficients, which are drawn from overall
coefficients. We also include a covariance structure among the intercept
and slope at each level. Formally, our hierarchical structure is as
follows, where we drop the intercept (0) and slope (1) subscripts and
instead refer to a vector of coefficients, \(\boldsymbol{\beta}\):

@@

We fit the model using a Bayesian approach, obtaining posterior
estimates of all unknowns via the No-U-Turn Hamiltonian Monte Carlo
sampler in Stan \citep{stan2016}. We used the R package `rstan'
\citep{rstan2016} to link R \citep{R2016} to Stan. We obtained samples
from the posterior distribution for all model parameters from four
parallel MCMC chains run for 10,000 iterations, saving every
\(10^{\text{th}}\) sample. Trace plots of all parameters were visually
inspected to ensure well-mixed chains and convergence. We also made sure
all scale reduction factors (\(\hat{R}\)) were less than 1.1.

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP
(mean of \(\beta_{1}\) = 0.41; Fig. 3), but the effect is associated
with high uncertainty, with credible intervals that overlap zero (80\%
BCI = -0.08, 0.86; 95\% BCI = -0.45, 1.13). Ecosystem functional
response was similar among treatments, with treatment level regressions
between soil moisture and log(ANPP) having similar intercepts and slopes
(Fig. 3A). There was also no evidence that the treatment effects became
more important over time because there was no directional trend in the
random year effects (Fig. 4).

@@

The similarity of ecosystem functional response (Fig. 3) and community
composition (Fig. 5) across treatments is surprising because grasslands
generally, and sagebrush steppe specifically, are considered
water-limited systems. Indeed, we expected ecosystem functional
response, community composition, or both to change with precipitation
treatment, landing us in any box of Fig. 1 \emph{except} the top left.
So, why did our treatments fail to induce ecosystem or community
responses? We can think of three reasons; two are limitations of our
study, and one is the life history traits of the species in our focal
communities. We first discuss the potential limitations of our study,
and then discuss the biological explanation.

First, it could be that our manipulations were not large enough to
induce a response. That is, maybe a 50\% decrease/increase in any given
year is not abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our manipulations \emph{should} have been
large enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation necessary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site. The 1\% quantile of
precipitation at our site is 110 mm, a 47\% reduction from the mean, and
the 99\% quantile is 414 mm, a 77\% increase from mean growing season
precipitation (Appendix 3). Thus, our drought treatment represented
extreme precipitation amounts, especially in years where ambient
precipitation was below average (Fig. 2A). The irrigation treatment may
have been too small, however.

Second, ANPP at our site may be influenced by factors beyond the window
of soil moisture we included in our statistical model. For example,
temperature can impact ANPP directly \citep{Epstein1997} and by
exacerbating the effects of soil moisture \citep{DeBoeck2011}.
Measurements of soil moisture likely contain a signal of temperature,
through its impact on evaporation and infiltration, but the measurements
will not capture the direct effect of temperature on metabolic and
physiological processes. Likewise, we did not redistribute snow across
our treatments in the winter, and snow melt may spur early spring
growth. These statistical issues of missing potentially important
covariates could explain the weak and uncertain relationship we observed
between soil moisture and ANPP.

@@

This research was supported by the Utah Agricultural Experiment Station,
Utah State University, and approved as journal paper number XXXX. The
research was also supported by the National Science Foundation, through
a Postdoctoral Research Fellowship in Biology and Mathematics to ATT
(DBI-1400370), a Graduate Research Fellowship to ARK, and grants
DEB-1353078 and DEB-1054040 to PBA.

@@

\begin{itemize}
  \item Andrew T. Tredennick collected data, analyzed the data, wrote the paper, prepared figures and/or tables, reviewed
drafts of the paper.
  \item Andrew R. Kleinhesselink conceived and designed the experiments, performed the experiments, collected data, reviewed drafts of the paper.
  \item Bret Taylor contributed reagents/materials/analysis tools, reviewed drafts of the paper.
  \item Peter B. Adler conceived and designed the experiments, performed the experiments, collected data, analyzed the data, reviewed
drafts of the paper.
\end{itemize}

\section{SUPPLEMENTAL INFORMATION}\label{supplemental-information}

\begin{description}
\item [Appendix 1.] Additional methods and information on estimating aboveground net primary productivity.
\item [Appendix 2.] Details of the hierarchical Bayesian model and Fig. A2-1.
\item [Appendix 3.] Details on analysis of precipitation historical range of variability and Fig. A3-1.
\end{description}

@@

\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis` tests whether treatments form unique clusters in multidimensial space; `betadisper` tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  21 &   2 & 1.02 & 0.42 \\ 
  2011 & betadisper &  21 &   2 & 2.23 & 0.14 \\ 
  2012 & adonis &  22 &   2 & 1.10 & 0.34 \\ 
  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
  2013 & adonis &  22 &   2 & 1.23 & 0.14 \\ 
  2013 & betadisper &  22 &   2 & 0.28 & 0.76 \\ 
  2014 & adonis &  22 &   2 & 0.95 & 0.54 \\ 
  2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
  2015 & adonis &  21 &   2 & 1.05 & 0.40 \\ 
  2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
  2016 & adonis &  21 &   2 & 1.07 & 0.33 \\ 
  2016 & betadisper &  21 &   2 & 0.50 & 0.62 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}
